Correlation between atom and cavity emissions in a driven Jaynes-Cummings system\\ \thanks{corresponding author e-mail: guillermopreisser@hotmail.com}

### 2019-09-17 research_paper/paper1.2.tex
Update paper1.2.tex
Add files via upload

%\begin{center}
%\begin{figure}[t!]
%\begin{center}
%\includegraphics[scale = 0.45]{image_paper222.png}
%\caption{\small{Calculation of correlation using 500 time intervals, fixing $\expval{n} \approx 1$, $\gamma =$ 1.0, $g = $ 0.1, and with  $\mathcal{E} =  \kappa |\langle n \rangle|[1 + 2g^2/\gamma \kappa]$.}} \label{corrxy}
%\end{center}  
%\end{figure}
%\end{center}
\section{Driven Jaynes-Cummings model}
We consider a system consisting of a two-level atom, coupled to a mode of a cavity. The mode is driven by a coherent field resonant with the atom and the mode of the cavity. In order to account for dissipation one may use a non-hermitian Hamiltonian to describe the system \cite{bla}, which will be useful at the moment of using quantum trajectory theory. Such Hamiltonian can be separated into hermitian and non-hermitian terms.
%\begin{equation}
%H = H_s - \frac{i\hbar}{2}\sum_m C^\dagger_n C_n .
%\end{equation}
The hermitian part of our Hamiltonian is given by $H_s$ which for the driven Jaynes-Cumming, under the rotating wave approximation, reads as
\begin{equation}
H_s = H_{at} + H_{cav} + H_{atcav} + H_{lascav}, \label{mainham}  
\end{equation}
where the atom and cavity terms are, respectively
\begin{subequations}
\begin{equation}
H_{at} = \tfrac{1}{2}\hbar \omega \sigma_z,    
\end{equation}
\begin{equation}
H_{cav} = \hbar \omega  a^\dagger a,  
\end{equation}
while the coupling between atom and cavity, and cavity and coherent field are
\begin{equation}
H_{atcav} = i\hbar g(a\sigma_+ - a^\dagger \sigma_-),    
\end{equation}
\begin{equation}
H_{lascav} = i\hbar \mathcal{E}(ae^{i\omega t} - a^\dagger e^{-i\omega t}),    
\end{equation}
\end{subequations}
where we introduced the coupling constant
\begin{equation}
g = \sqrt{\frac{\omega d^2}{2\hbar \epsilon_0 V_Q}},
\end{equation}
with the mode volume $V_Q$, dipolar coupling constant $d$, and $\epsilon_0$ is the permitivity of free space. $\mathcal{E}$ is a term proportional to the coherent field amplitude.
As for the non-hermitian part of our Hamiltonian, we have
\begin{equation} \label{nh}
H_{nh} = H_{dec} + H_{leak},
\end{equation}
where the terms corresponding to the spontaneous emission of the atom and the cavitty losses are
\begin{subequations}
\begin{equation}
H_{dec} = - i\hbar\frac{\gamma}{2}\sigma_+\sigma_-,
\end{equation}
\begin{equation}
H_{leak} = - i\hbar\kappa a^\dagger a.
\end{equation}
\end{subequations}
The annihilation operator $a$ of a single photon in the cavity, and the creation operator $a^\dagger$ obey $[a, a^\dagger] = 1$, while the operators $\sigma_-, \sigma_+$  account for the atomic lowering and raising operators, respectively, and the operator $\sigma_z$ accounts for the population difference between the atomic levels. These operators obey $[\sigma_+, \sigma_-] = \sigma_z, \ [\sigma_\pm, \sigma_z] = \mp 2\sigma_\pm$. $\kappa$ and $\gamma$ are the cavity loss and spontaneous decay rate.\\ %With the Hamiltonian Eq. \eqref{mainham} we can write down the master equation for the driven Jaynes-Cummings model in its Lindblad form
%\begin{equation}
%\frac{\partial \rho}{\partial t} = -\frac{i}{\hbar} [H(t), \rho]  + \mathcal{L}_\kappa \rho + \mathcal{L}_{\gamma} \rho,
%\end{equation}
%where 
%\begin{equation}
%\mathcal{L}_\kappa  \rho =  \mathcal{D}[\sqrt{2\kappa}a] \rho, 
%\end{equation}
%and
%\begin{equation}
%\mathcal{L}_\gamma \rho = \ \mathcal{D}[\sqrt{\gamma}\sigma_-] \rho,
%\end{equation}
%where $\mathcal{D}$ is a superoperator that will act on the density matrix as
%\begin{equation}
%\mathcal{D}[C] = \frac{1}{2}[2C\rho C^\dagger - C^\dagger C \rho - \rho C^\dagger   C],
%\end{equation}
%where we identify $C$ as the jump operators which will be the ones that will cause the dissipation in the system when we work using quantum trajectory theory.
%\begin{figure*}[t!]
%\includegraphics[scale = 0.8]{phases.png}
%\end{figure*}
%\begin{SCfigure*}[0.9][!h] \label{figsis}

@@

\section{Estimation of spontaneous emissions}
In this section we obtain estimations for the number of spontaneous emissions by measuring stimulated emission in a cavity mode in the driven Jaynes-Cummings model, using quantum trajectory theory.
\subsection{Probability distributions of emissions}
In order to get the desired estimations, first we must obtain probability distributions that relate the number of spontaneous emissions of the atom with the emissions measured from the cavity. We achieve this by making a large number of quantum  trajectories from which we keep a record of the number of spontaneous emissions and cavity losses, relating the number of spontaneous emissions for each number of cavity loss and making a probability distribution out of the emission's register of the trajectories. \\

In order to achieve this, following the procedure from quantum trajectory theory as it is explained in \cite{bla}. Starting with a state which represents a ground state of the atom with no photons in the cavity, $\ket{\phi(t = 0)} = \ket{\downarrow} \times \ket{0}$, we evolve it with the Schrödinger equation through some time $\delta t$, using a non-hermitian Hamiltonian of the form
\begin{equation}
H = H_s - \frac{i\hbar}{2}\sum_n C^\dagger_n C_n,
\end{equation}
where the operators $C_n$ are known as collapse operators.
Using a non-hermitian Hamiltonian will have as a consequence that the probability won't be conserved \cite{Sakurai:1167961}. The evolution of the state through the schrödinger equation will be interrupted by a quantum jump – an abrupt change on our state in which a photoemission has ocurred. The probability for such quantum jump to happen will be given by the collapse probability
\begin{equation} \label{probcol}
\delta p = \sum_n \delta p_n,
\end{equation}
where 
\begin{equation}
\delta p_n = \delta t \bra{\phi (t)} C^\dagger_n C_n \ket{\psi (t)}.
\end{equation}
We use these collapse probabilities to simulate the quantum trajectories, where we compare the collapse probability of eq. \eqref{probcol} with a random number $\epsilon$, uniformly distributed between zero and one. In the case the condition $\delta p > \epsilon$ is fulfilled we simulate that a quantum jump has happened, by acting with one of the collapse operators. From the non-hermitian Hamiltonian of our system \eqref{nh} we are able to identify the corresponding collapse operators which are
\begin{figure}[!t] 
\centering
\includegraphics[scale = 0.5]{distributioneng.png}
\caption{\small{Probability distribution of spontaneous emissions for different number of cavity losses in a driven Jaynes-cummings model, doing $100,000$ quantum trajectory simulations, with parameters $\mathcal{E}  = \kappa = \gamma = g = 1$.}} \label{probdiss}
\end{figure}
\begin{subequations}
\begin{equation}
C_1 = \sqrt{\gamma}\sigma_-,
\end{equation}
\begin{equation}
C_2 = \sqrt{2\kappa}a.
\end{equation}
\end{subequations}
Using equation \eqref{probcol} one can easily identify the corresponding collapse probabilities of the system. Having identified the collapse operators and the collapse probability, we simulate quantum trajectories, in which we keep count of the number of quantum jumps for each collapse operator, relating the number of quantum jumps associated to spontaneous emissions for each number of quantum jumps associated to cavity emissions until we reach a maximum number of cavity emissions, which we establish. The procedure of register of emissions is summarized in algorithm \ref{alg1}. \\

>>>>

%\begin{center}
%\begin{figure}[t!]
%\begin{center}
%\includegraphics[scale = 0.45]{image_paper222.png}
%\caption{\small{Calculation of correlation using 500 time intervals, fixing $\expval{n} \approx 1$, $\gamma =$ 1.0, $g = $ 0.1, and with  $\mathcal{E} =  \kappa |\langle n \rangle|[1 + 2g^2/\gamma \kappa]$.}} \label{corrxy}
%\end{center}  
%\end{figure}
%\end{center}
\section{Driven Jaynes-Cummings model}
We consider a system consisting of a two-level atom, coupled to a mode of a cavity. The mode is driven by a coherent field resonant with the atom and the mode of the cavity. In order to account for dissipation one may use a non-hermitian Hamiltonian to describe the system \cite{bla}. Using this description for the system will turn useful at the moment of using quantum trajectory theory. Such Hamiltonian can be separated into hermitian and non-hermitian terms.
%\begin{equation}
%H = H_s - \frac{i\hbar}{2}\sum_m C^\dagger_n C_n .
%\end{equation}
The hermitian part of our Hamiltonian is given by $H_s$ which for the driven Jaynes-Cumming, under the rotating wave approximation, reads as
\begin{equation}
H_s = H_{at} + H_{cav} + H_{atcav} + H_{lascav}, \label{mainham}  
\end{equation}
where the atom and cavity terms are, respectively
\begin{subequations}
\begin{equation}
H_{at} = \tfrac{1}{2}\hbar \omega \sigma_z,    
\end{equation}
\begin{equation}
H_{cav} = \hbar \omega  a^\dagger a,  
\end{equation}
while the coupling between atom and cavity, and cavity and coherent field are
\begin{equation}
H_{atcav} = i\hbar g(a\sigma_+ - a^\dagger \sigma_-),    
\end{equation}
\begin{equation}
H_{lascav} = i\hbar \mathcal{E}(ae^{i\omega t} - a^\dagger e^{-i\omega t}),    
\end{equation}
\end{subequations}
where we introduced the coupling constant
\begin{equation}
g = \sqrt{\frac{\omega d^2}{2\hbar \epsilon_0 V_Q}},
\end{equation}
with the mode volume $V_Q$, dipolar coupling constant $d$, and $\epsilon_0$ is the permitivity of free space. $\mathcal{E}$ is a term proportional to the coherent field amplitude.
As for the non-hermitian part of our Hamiltonian, we have
\begin{equation} \label{nh}
H_{nh} = H_{dec} + H_{leak},
\end{equation}
where the terms corresponding to the spontaneous emission of the atom and the cavitty losses are
\begin{subequations}
\begin{equation}
H_{dec} = - i\hbar\frac{\gamma}{2}\sigma_+\sigma_-,
\end{equation}
\begin{equation}
H_{leak} = - i\hbar\kappa a^\dagger a.
\end{equation}
\end{subequations}
The annihilation operator $a$ of a single photon in the cavity, and the creation operator $a^\dagger$ obey $[a, a^\dagger] = 1$. The operators $\sigma_-, \sigma_+$  account for the atomic lowering and raising operators, respectively, and the operator $\sigma_z$ accounts for the population difference between the atomic levels. These operators obey $[\sigma_+, \sigma_-] = \sigma_z, \ [\sigma_\pm, \sigma_z] = \mp 2\sigma_\pm$. $\kappa$ and $\gamma$ are the cavity loss and spontaneous decay rate.\\ %With the Hamiltonian Eq. \eqref{mainham} we can write down the master equation for the driven Jaynes-Cummings model in its Lindblad form
%\begin{equation}
%\frac{\partial \rho}{\partial t} = -\frac{i}{\hbar} [H(t), \rho]  + \mathcal{L}_\kappa \rho + \mathcal{L}_{\gamma} \rho,
%\end{equation}
%where 
%\begin{equation}
%\mathcal{L}_\kappa  \rho =  \mathcal{D}[\sqrt{2\kappa}a] \rho, 
%\end{equation}
%and
%\begin{equation}
%\mathcal{L}_\gamma \rho = \ \mathcal{D}[\sqrt{\gamma}\sigma_-] \rho,
%\end{equation}
%where $\mathcal{D}$ is a superoperator that will act on the density matrix as
%\begin{equation}
%\mathcal{D}[C] = \frac{1}{2}[2C\rho C^\dagger - C^\dagger C \rho - \rho C^\dagger   C],
%\end{equation}
%where we identify $C$ as the jump operators which will be the ones that will cause the dissipation in the system when we work using quantum trajectory theory.
%\begin{figure*}[t!]
%\includegraphics[scale = 0.8]{phases.png}
%\end{figure*}
%\begin{SCfigure*}[0.9][!h] \label{figsis}

@@

\section{Estimation of spontaneous emissions}
In this section we obtain estimations for the number of spontaneous emissions by measuring stimulated emission in a cavity mode in the driven Jaynes-Cummings model, using quantum trajectory theory.
\subsection{Probability distributions of emissions}
In order to get the desired estimations, first we must obtain probability distributions that relate the number of spontaneous emissions of the atom with the emissions measured from the cavity. We achieve this by making a large number of quantum  trajectories from which we keep a record of the number of spontaneous emissions and cavity losses, relating the number of spontaneous emissions for each number of cavity loss and making a probability distribution out of the emission's register of the trajectories.

In order to achieve this, following the procedure from quantum trajectory theory as it is explained in \cite{bla}. Starting with a state which represents a ground state of the atom with no photons in the cavity, $\ket{\phi(t = 0)} = \ket{\downarrow} \times \ket{0}$, we evolve it with the Schrödinger equation through some time $\delta t$, using a non-hermitian Hamiltonian of the form
\begin{equation}
H = H_s - \frac{i\hbar}{2}\sum_n C^\dagger_n C_n,
\end{equation}
where the operators $C_n$ are known as collapse operators.
Using a non-hermitian Hamiltonian will have as a consequence that the probability won't be conserved \cite{Sakurai:1167961}. The evolution of the state through the schrödinger equation will be interrupted by a quantum jump – an abrupt change on our state in which a photoemission has ocurred. The probability for such quantum jump to happen will be given by the collapse probability
\begin{equation} \label{probcol}
\delta p = \sum_n \delta p_n,
\end{equation}
where 
\begin{equation} \label{relprob}
\delta p_n = \delta t \bra{\phi (t)} C^\dagger_n C_n \ket{\psi (t)}.
\end{equation}
We use these collapse probabilities to simulate the quantum trajectories, where we compare the collapse probability of eq. \eqref{probcol} with a random number $\epsilon$, uniformly distributed between zero and one. In the case the condition $\delta p > \epsilon$ is fulfilled we simulate that a quantum jump has happened, by acting with one of the collapse operators. The operator with which we'll act on the state will be given according to the relative probabilities of eq. \eqref{relprob}. From the non-hermitian Hamiltonian of our system \eqref{nh} we are able to identify the corresponding collapse operators which are
\begin{figure}[!t] 
\centering
\includegraphics[scale = 0.5]{distributioneng.png}
\caption{\small{Probability distribution of spontaneous emissions for different number of cavity losses in a driven Jaynes-cummings model, doing $100,000$ quantum trajectory simulations, with parameters $\mathcal{E}  = \kappa = \gamma = g = 1$.}} \label{probdiss}
\end{figure}
\begin{subequations}
\begin{equation}
C_1 = \sqrt{\gamma}\sigma_-,
\end{equation}
\begin{equation}
C_2 = \sqrt{2\kappa}a.
\end{equation}
\end{subequations}
Using equation \eqref{probcol} one can easily identify the corresponding collapse probabilities of the system. Having identified the collapse operators and the collapse probability, we simulate quantum trajectories, in which we keep count of the number of quantum jumps for each collapse operator, relating the number of quantum jumps associated to spontaneous emissions for each number of quantum jumps associated to cavity emissions until we reach a maximum number of cavity emissions, which we establish. The procedure of register of emissions is summarized in algorithm \ref{alg1}.
